---
title: "Porous model example run"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Porous model example run}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Intro
This poorly written vignette contains some working example of how to run the model in its current state.  
The code is organized in a series of functions:


```{r,echo=FALSE, fig.height = 3, fig.width = 7}
library(DiagrammeR)
DiagrammeR::grViz("digraph {

graph [layout = dot, rankdir = LR]
# define the global styles of the nodes. We can override these in box if we wish
node [shape = rectangle, style = filled, fillcolor = Linen]

Delta_z [label =  'Delta_z function', fillcolor = cornflowerblue]
phi_mat [label = 'phi_mat function', fillcolor = darkseagreen3]
phi_mic [label = 'phi_mic function', fillcolor = bisque2]
f_text_mic_func [label = 'f_text_mic_func function', fillcolor = mediumpurple1]
Porous [label = 'Porous function', fillcolor = darkorange]
pore_frac [label= 'pore_frac function', fillcolor = goldenrod2]
Msc [label= 'Msc function', fillcolor = yellowgreen]
f_som [label= 'f_som function', fillcolor = hotpink]
gamma_b [label= 'gamma_b function', fillcolor = seagreen2]
run_Porous [label= 'run_Porous function', fillcolor = peachpuff]
Delta_z_2 [label =  'Delta_z function, outside run', fillcolor = cornflowerblue]

# edge definitions with the node IDs
Delta_z -> phi_mat -> pore_frac
Delta_z -> phi_mic -> pore_frac
f_text_mic_func -> phi_mic
pore_frac -> Porous
Porous -> run_Porous
Msc -> f_som -> run_Porous
Msc -> gamma_b -> run_Porous
Porous -> Delta_z_2 -> run_Porous

}")
```



## Example: how to to run the main model function (written within the SoilR model development framework)

```{r setup}
library(Porous)
library(SoilR) #this should load automatically anyway since it is required by Porous
```

First we create the model object:
```{r}
 modelObject<-Porous()
```
This creates a model with the default parameters. The parameter values can also be specified by the user (when one value is not specified the parameter will have default value), such as:

```{r}
modelObject<-Porous(ky=0.8, ko=0.00605,
                    kmix=0.9,
                    e=0.13,
                    Im=1.1, Ir=0.5,
                    F_prot=0.1,
                    phi_mac=0.2,
                    phi_min=1,
                    clay=0.2,
                    Delta_z_min=20,
                    gamma_o=1.2,
                    f_agg=3)

```
Many values can be omitted, and the function will use the defaults. In particular `f_text_mic `, if omitted, will be calculated internally with a specific function, `f_text_mic_func()`.

`SoilR` has an internal function designed to plot any model (but fluxes do not always seem consistent in the representation):

```{r, fig.height = 7, fig.width = 7}
plotPoolGraph(modelObject)
```

Now we can run the model, for which we need to initialize the starting values
```{r}
init<-c(My_mes=1, Mo_mes=10,My_mic=0.6, Mo_mic=3)
times<-seq(0,20,by=0.1)
```

We can then run the model with the specific `SoilR` function:
```{r}
modrun0<-Model_by_PoolNames(smod=modelObject, times=times, initialValues=init)
```

And, again relying on  the specific `SoilR` function, we can extract the C stocks and respiration fluxes:

```{r}
Stocks<-getC(modrun0)
Resp<-getReleaseFlux(modrun0)
```
Let's see what object we created (by plotting just the first rows of it):
```{r}
head(Resp)
```

Everything seems in order and we can proceed to plot:
```{r fig1, dpi = 72, fig.height = 8, fig.width = 8}
par(mfrow=c(2,1), mar=c(4,4,0,1))
matplot(times, Stocks, type="l", lty=1, col=1:4, xlab=" ", ylab="Pool contents", bty="n")
legend("bottomright", c("My_mes", "Mo_mes", "My_mic", "Mo_mic"), lty=1, col=1:4, bty="n")
matplot(times, Resp,  type="l", lty=1, col=1:2, xlab="Time", ylab="Respiration", bty="n")
```


It is also possible to calculate the evolution of soil bulk densities and soil C concentration based on the simulated C stocks.  
The soil bulk density relies on the function `gamma_b`:  
```{r}
f_som(My_mic=Stocks[1:20,1],
      Mo_mic=Stocks[1:20,2],
      My_mes=Stocks[1:20,3],
      Mo_mes=Stocks[1:20,4],
      Delta_z_min=20,
      phi_min=0.35,
      gamma_m=1.1)
```

While the soil C concentration on the function `f_som`:  
```{r}
gamma_b(My_mic=Stocks[1:20,1],
        Mo_mic=Stocks[1:20,2],
        My_mes=Stocks[1:20,3],
        Mo_mes=Stocks[1:20,4],
        Delta_z_min=20,
        phi_min=1,
        gamma_o=1.2,
        gamma_m=1.1,
        f_agg=3,
        phi_mac=0.2)
```


## Running the model with the wrapper (more comfortable, less transparent)
All the above functions can be run with a wrapper, `run_Porous`, that takes care of running the simulation and calculating bulk density, C concentration and layer thickness evolution:

```{r}
simulation<-run_Porous(ky=0.8, ko=0.04,
                     kmix=0.02,
                     e=0.15,
                     Im=0.08, Ir=0.048,
                     F_prot=0.1,
                     phi_mac=0.04,
                     clay=0.2,
                     Delta_z_min=14.2,
                     gamma_o=1.2,
                     gamma_m=2.7,
                     proportion=NULL,
                     phi_min=0.35,
                     f_text_mic=0.5069,
                     f_agg=3,
                     init=c(My_mes=0.132, Mo_mes=0.507, My_mic=0.284, Mo_mic=0.579),
                     sim_length=20,
                     sim_steps=0.1)

simulation$`mass balance`

results<-simulation$results

head(results)
```
The wrapper first runs a C stocks simulation (including with variable layer thickness calculation) and then it uses the C stocks from the simulation to calculate again the layer thickness and calculate for the first time SOC concentration and soil bulk density to include these variables in the results. This is done because it is not possible (or I do not know the way) to export the internal calculations from the SoilR model.  
Please notice that the above code is overriding the `f_text_mic` calculation.


Results can be accessed as plotted as a data frame:  
```{r fig2, dpi = 72, fig.height = 5, fig.width = 8}
range_stocks<-range(results[,2:5])
plot(results$time, results$My_mes.stocks, ylim=c(-2,range_stocks[2]), type="l", col=1, xlab="Years", ylab="C stocks")
lines(results$time, results$Mo_mes.stocks, ylim=range_stocks, type="l", col=2)
lines(results$time, results$My_mic.stocks, ylim=range_stocks, type="l", col=3)
lines(results$time, results$Mo_mic.stocks, ylim=range_stocks, type="l", col=4)
legend("bottomright", c("My_mes", "Mo_mes", "My_mic", "Mo_mic"), lty=1, col=1:4, bty="n")
```

The `run_Porous`: function, on top of the simulation of C stocks from the main model function `porous`, gives also the simulated evolution of bulk density, the C concentration and layer thickness:
```{r fig3, dpi = 72, fig.height = 12, fig.width = 8}
par(mfrow=c(3,1))
plot(results$time, results$gamma_b_sim,  type="l", col=6, xlab="Years", ylab="Bulk density")
plot(results$time, results$f_som_sim,  type="l", col=7, xlab="Years", ylab="C concentration")
plot(results$time, results$Delta_z_sim,  type="l", col=5, xlab="Years", ylab="Delta Z (cm)")
```

  
The package contains an alternative implementation of the same model, directly written around the `deSolve` package (but wrapped to have the same syntax). The following is an example of a model run at equilibrium (so the pools are supposed to be flat):

```{r}
simulation_deSolve<-run_Porous_deSolve(ky=0.8, ko=0.04,
                                        kmix=0.02,
                                        e=0.15,
                                        Im=0.08, Ir=0.048,
                                        F_prot=0.1,
                                        phi_mac=0.04,
                                        clay=0.2,
                                        Delta_z_min=14.2,
                                        gamma_o=1.2,
                                        gamma_m=2.7,
                                        proportion=NULL,
                                        phi_min=0.35,
                                        f_text_mic=0.5069,
                                        f_agg=3,
                                        init=c(My_mes=0.132, Mo_mes=0.507, My_mic=0.284, Mo_mic=0.579),
                                        sim_length=30,
                                       sim_steps=0.1)

results_deSolve<-simulation_deSolve$results

head(results_deSolve)
```



```{r fig4, dpi = 72, fig.height = 5, fig.width = 8}
range_stocks<-range(results_deSolve[,2:5])
plot(results_deSolve$time, results_deSolve$My_mes.stocks, ylim=c(-2,range_stocks[2]), type="l", col=1)
lines(results_deSolve$time, results_deSolve$Mo_mes.stocks, col=2)
lines(results_deSolve$time, results_deSolve$My_mic.stocks, col=3)
lines(results_deSolve$time, results_deSolve$Mo_mic.stocks, col=4)
```
